---
title: "Untitled"
output: github_document
always_allow_html: true

#Bhan, Lam (2022) [Nanyang Technological University]
#ORCID: 0000-0001-5193-6560
#GitHub: https://github.com/bhanlam/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)

library(pander)
library(dataverse)
library(tibble)
library(readxl)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(forcats)
library(skimr)
library(janitor)
library(rstatix)
library(ggplot2)
```

This repository contains the code accompanying the publication "Do uHear? Validation of uHear App for Preliminary Screening of Hearing Ability in Soundscape Studies". This paper is hosted on [arXiv]() and the raw dataset is hosted at [10.21979/N9/0NE37R](https://doi.org/10.21979/N9/0NE37R).

## Environment

The code has been tested on the following platform.

```{r session, message=FALSE, echo=FALSE}
pander(sessionInfo())
```

## Data preparation

Load data from excel

```{r dataloader, message=FALSE}
#load data into dataframe
dataset <- tibble::as_tibble(read_excel("Hearing test result.xlsx",skip = 0))


#prepare data in dataframe for analysis
cols2fac <- colnames(dataset)[colnames(dataset)!="Age"]
facOrder <- c(seq(-10,70,5),"NaN") #order factors
cleandata <- dataset %>% 
        filter(complete.cases(.)) %>% #remove incomplete cases
        mutate_at(cols2fac,factor) %>% #convert variable to factor 
        mutate_at(cols2fac[5:15],~fct_relevel(.,facOrder))

#tidy data to wide form
widedata <- cleandata %>%
        pivot_wider(names_from=c(Type,Ear),values_from = c("0.125":"8"))

```

## Exploratory analysis

### Demographics

```{r demographics}
demo <- cleandata %>% #extract demographics only
        select(c("Participant ID","Gender","Age")) %>%
        distinct(`Participant ID`,.keep_all = TRUE)
skim(demo)
get_summary_stats(demo) #summary stats for age
demo %>% tabyl(Gender) #gender count w perc

#generate latex table

```
### Pure tone audiometry summary statistics

Summarize proportions across dB levels for each frequency

```{r ptaSummary}

t<-cleandata %>% 
        filter(Type=="Audiometer") %>%
        tabyl(`0.125`,Ear) %>%
        adorn_totals(where = "row") %>%             # add a total row
        adorn_percentages(denominator = "col") %>%  # convert to proportions
        adorn_pct_formatting() %>%                  # convert to percents
        adorn_ns(position = "front") %>%            # display as: "count (percent)"
        adorn_title(                                # adjust titles
        row_name = "125 Hz",
        col_name = "Ear")

```


### Bland-Altman

Bland-Altman plots to visually compare differences between audiometer and mobile app

UHear
- 0.5 kHz
- 1 kHz
- 2 kHz
- 4 kHz
- 6 kHz

J. Martin Bland, D. Altman, STATISTICAL METHODS FOR ASSESSING AGREEMENT BETWEEN TWO METHODS OF CLINICAL MEASUREMENT, Lancet. 327 (1986) 307–310. https://doi.org/10.1016/S0140-6736(86)90837-8.

```{r blaalt}

uHfreq<-c("0.5","1","2","4","6")
comparedf<-widedata %>% select("Participant ID","Gender","Age")

#compute mean and difference for each frequency across left and right
for (freq in uHfreq){
        comparedf<-cbind(comparedf,dataset %>% 
                filter(complete.cases(.)) %>% #remove incomplete cases
                pivot_wider(names_from=c(Type,Ear),
                            values_from = c("0.125":"8")) %>% #expanding Type and Ear
                select(matches(paste0("^",freq,"_"))) %>% #0.5 kHz
                #add freq as colname to reflect in variable
                mutate(meanL=rowMeans(select(., contains("Left")),na.rm=T), 
                       meanR=rowMeans(select(., contains("Right")),na.rm=T),
                       diffL=(select(., contains("Left") & contains("Audiometer"))-
                               select(., contains("Left") & contains("uHear")))[,],
                       diffR=(select(., contains("Right") & contains("Audiometer"))-
                               select(., contains("Right") & contains("uHear")))[,]) %>%
                dplyr::rename(!!paste0("meanL",freq):=`meanL`) %>%
                dplyr::rename(!!paste0("meanR",freq):=`meanR`) %>%
                dplyr::rename(!!paste0("diffL",freq):=`diffL`) %>%
                dplyr::rename(!!paste0("diffR",freq):=`diffR`)) 
}

ggplot(comparedf, aes(x = meanL0.5, y = diffL0.5)) +
  geom_point(size=2) +
  geom_hline(yintercept = mean(comparedf$diffL0.5)) +
  geom_hline(yintercept = mean(comparedf$diffL0.5) 
             - 1.96*sd(comparedf$diffL0.5) , color = "red", linetype="dashed") +
  geom_hline(yintercept = mean(comparedf$diffL0.5) 
             + 1.96*sd(comparedf$diffL0.5), color = "red", linetype="dashed") +
  ggtitle("Bland-Altman Plot") +
  ylab("Difference Between Instruments") +
  xlab("Average")+theme_bw()

```

## Statistical analysis





